Exposure Tradeoff Thing
=======================

This is me playing with numbers. It is not real science. It is certainly
not real epidemiology, and no one should take it seriously or follow my 
advice. I did this to help myself think about the risks associated with 
various outdoor activities. In particular, the difference between a
"distance date" at a park or in a yard with a single friend versus going 
for a solo walk or bike ride on a populated trail.

You can find the code [here](./Tradeoffs.Rmd).

## Setup
```{r, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
theme_set(theme_bw())
```

## Description

### Parameters:

* `prev` is the prevalence of infectious people in the population. It 
ranges from 1/10 to 1/5000
* `er` is the exposure ratio. It can be thought of as (crowd density * 
exposure time)/exposure threshold. So, for example, if the threshold amount
of exposure required to become infected is 1 hour and you spend 1 hour in 
the presence of sets of three people _simultaneously_, or 3 hours in the 
presence of 1 person _at a time_, then your exposure ratio would be 3. 
It ranges from 1 to 40. 
* `n` is the total number of people you are exposed to during your exposure
time. For a given exposure ratio, the higher `n` is, the less time you 
spend with each person. I have `n` ranging from 1 to 500.

Here are some scenarios to elucidate the meanings of the `er` and `n` 
parameters:

* If it takes 1 hour of exposure to become infected, and I spend 1 hour
with 1 person, that gives me an **exposure ratio of 1** and **n of 1**
* If it takes 1 hour of exposure to become infected and
I have 4 15-minute meetings back-to-back with 1 person each, then I have an
**exposure ratio of 1** again, but an **n of 4**
    * If my 15-minute meetings were with 2 people each then I would have an
    **exposure ratio of 2** and an **n of 8**
* If it takes 1 hour of exposure to become infected, and I spend 3 hours at
a protest where I am only ever exposed to 5 people at a time, but I move 
around a lot so the total number of people I am exposed to is 50 then I
will have an **exposure ratio of 15** (3 hours x 5 people) and an 
**n of 50**
* If it takes 15 minutes of exposure to become infected and I attend an
hour-long meeting with 10 people then I will have an 
**exposure ratio of 40** and an **n of 10**.
* If it takes 3 hours to get infected and I spend 2 hours continuously in 
the presence of 1 person at a time, I will have an 
**exposure ratio of 2/3**
    * Any time exposure ratio is less than 1, the probability of infection 
    is 0, so that situation is excluded from the calculations below
        * Keep in mind that this is actually false. Infection probability
        would be better modeled as an exponential function of exposure 
        ratio, not a step function. But now we're veering a little close to
        real science.

### Calculations:

I iterate through all the parameter values mentioned above. I also 
set the probability of infection when you're exposed to a single
individual as equal to the prevalence.

In the case of exposure to a single individual, the probability of becoming
infected (assuming `er` is at least 1) is equal to the prevalence of the
infection in the population.

In the case of multiple shorter exposures to different sets of people, I 
calculate the threshold number of people required to give
a threshold exposure as `n_thresh = n/er`. That is, if my exposure ratio is
2 then I need at least half of the people who I encounter to be infected in
order to become infected myself. That is the same as saying I need at least
`n/er` people to be infected in order to reach an _effective_ exposure 
ratio of 1.

The probability of at least `n_thresh` hits in a crowd of size `n` is given
by the area under the region of the probability density function of a
binomial distribution with size n and p = prevalence where hits are greater
than or equal to `n_thresh`.

```{r}
# prevalence of infections in the population
prevs = c(1/10,1/100,1/500,1/1000,1/5000)

# The total number of people you are exposed to during your exposure. 
# The higher this number, the less time you spend with each individual
ns = seq(500)

# Exposure ratios
ers = seq(40)

nrows = length(prevs)*length(ns)*length(ers)
res_mat = matrix(nrow = nrows, ncol = 6)
colnames(res_mat) = c('Prevalence', 'ExposureRatio',
                      'N','NThresh','PSingle',
                      'PCrowd')
i = 1
for (prev in prevs){
    for (n in ns){
        for (er in ers){
            p_single = prev
            n_thresh = ceiling(n/er)
            p_crowd = pbinom(n_thresh - 1, n, prev,
                             lower.tail = FALSE)
            res_v = c(prev, er,
                      n, n_thresh, p_single, p_crowd)
            res_mat[i,] = res_v
            i = i + 1
        }
    }
}
res_mat = na.omit(res_mat)
#dim(res_mat)
df = (res_mat 
      %>% data.frame())
#head(df)
#summary(df)

#df %>% filter(PCrowd > PSingle, N > 1) %>% summary()

df_long = (df
      %>% gather(P_type, Probability, PSingle, PCrowd))
#head(df_long)
df_plt = (df_long
          %>% filter(ExposureRatio %in% c(1,2,10,25,40))
          %>% mutate_at(c('ExposureRatio','Prevalence'),as.factor))

plt = ggplot(df_plt, aes(N, Probability)) +
    geom_line(aes(colour = Prevalence, linetype = P_type)) +
    facet_wrap(~ExposureRatio) +
    scale_y_continuous(trans = 'log10', limits = c(1e-15,1)) +
    scale_colour_brewer(palette = 'Dark2')
plt
pdf('tradeoffs.pdf')
plt
dev.off()
```

## Some facts:

* The jaggedness of the lines is caused by the fact that n_thresh is
required to be an integer, so there is rounding happening.
* I limited the plots to regions where the probability of exposure was 
at least 10^-15. 
* When exposure ratio is small, rotating through people drops your
risk of infection _fast_. The higher your exposure ratio, the slower that
effect is, and if prevalence is high enough, increasing the number of 
people you are exposed to actually _increases_ your chances of exposure.

That box where the exposure ratio is 10 is making me suspicious. Let's see:

```{r}
ers = c(9,10,99,100,499,500,999,1000,4999,5000)
ns = seq(5000)

nrows = length(prevs)*length(ns)*length(ers)
res_mat = matrix(nrow = nrows, ncol = 6)
colnames(res_mat) = c('Prevalence', 'ExposureRatio',
                      'N','NThresh','PSingle',
                      'PCrowd')
i = 1
for (prev in prevs){
    for (n in ns){
        for (er in ers){
            p_single = prev
            n_thresh = ceiling(n/er)
            p_crowd = pbinom(n_thresh - 1, n, prev,
                             lower.tail = FALSE)
            res_v = c(prev, er,
                      n, n_thresh, p_single, p_crowd)
            res_mat[i,] = res_v
            i = i + 1
        }
    }
}

res_mat = na.omit(res_mat)
df = (res_mat 
      %>% data.frame())

df_long = (df
      %>% gather(P_type, Probability, PSingle, PCrowd))
#head(df_long)
df_plt = (df_long
          %>% mutate_at(c('ExposureRatio','Prevalence'),as.factor))

plt = ggplot(df_plt, aes(N, Probability)) +
    geom_line(aes(colour = Prevalence, linetype = P_type)) +
    facet_wrap(~ExposureRatio) +
    scale_y_continuous(trans = 'log10', limits = c(1e-15,1)) +
    scale_colour_brewer(palette = 'Dark2')
plt
pdf('tradeoffs_er_v_prev.pdf', width = 10)
plt
dev.off()

```

So, there's actual math to be done here but it _seems_ like PCrowd will
eventually fall below PSingle as long as prevalence is less than 
1/exposure ratio, and will approach 1 if pravalence is 
greater than 1/exposure ratio, and I'm not sure which it does if the
two are equal but whatever it does, it does it slowly.

```{r}
ns = seq(5e9,10e9,1e6)
nrows = length(prevs)*length(ns)*length(ers)
res_mat = matrix(nrow = nrows, ncol = 6)
colnames(res_mat) = c('Prevalence', 'ExposureRatio',
                      'N','NThresh','PSingle',
                      'PCrowd')
i = 1
for (prev in prevs){
    for (n in ns){
        for (er in ers){
            p_single = prev
            n_thresh = ceiling(n/er)
            p_crowd = pbinom(n_thresh - 1, n, prev,
                             lower.tail = FALSE)
            res_v = c(prev, er,
                      n, n_thresh, p_single, p_crowd)
            res_mat[i,] = res_v
            i = i + 1
        }
    }
}

res_mat = na.omit(res_mat)
df = (res_mat 
      %>% data.frame())

df_long = (df
      %>% gather(P_type, Probability, PSingle, PCrowd))
#head(df_long)
df_plt = (df_long
          %>% mutate_at(c('ExposureRatio','Prevalence'),as.factor))

plt = ggplot(df_plt, aes(N, Probability)) +
    geom_line(aes(colour = Prevalence, linetype = P_type)) +
    facet_wrap(~ExposureRatio) +
    scale_y_continuous(trans = 'log10', limits = c(1e-15,1)) +
    scale_colour_brewer(palette = 'Dark2')
plt
pdf('tradeoffs_er_v_prev_long.pdf', width = 10)
plt
dev.off()
```

Like, really slowly. Maybe it converges on a value less than 1 but greater 
than PSingle?

```{r}
summary((df_plt
    %>% filter(Prevalence == 0.1, ExposureRatio == 10, 
               P_type == 'PCrowd'))$Probability)
summary((df_plt
    %>% filter(Prevalence == 0.01, ExposureRatio == 100, 
               P_type == 'PCrowd'))$Probability)
summary((df_plt
    %>% filter(Prevalence == 0.002, ExposureRatio == 500, 
               P_type == 'PCrowd'))$Probability)
```

K so it's 1/2. 

I bet the math is probably not hard if you know how to do it. I
do not know how to do it.

## Conclusions

Possibly counterintuitively:

* Walking/biking along a moderately crowded path where you are constantly 
passing people is safer than spending that same amount of time sitting in 
an equivalent proximity to a single random person.

* If you are at a crowded outdoor event (say, a protest) it is probably 
safer to move around and vary which people you are in proximity to than to 
spend a lot of time in proximity to the same person or people.

* These guidelines only work if prevalence is fairly low relative to the
number of people you're encountering.

pdfs of plots are [here](./tradeoffs.pdf), 
[here](./tradeoffs_er_v_prev.pdf), and [here](./tradeoffs_er_v_prev_long.pdf)